Paper figures

### 2019-01-06 da_test.Rmd
new figures. Fixed family levels boxplots

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])

sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

unass_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(Domain == 'Unassigned') %>%
                select(OTU)

unknown_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Genus)) %>%
                select(OTU)

unknown_family <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Family)) %>%
                select(OTU)

tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
tax_table(gg_topi)[unknown_family$OTU, 'Family'] <- 'Unknown'

@@

gg_topi.family <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
```

@@

### Dotplot log2FC vs Family
```{r}
res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),-7])) %>%
  ggplot(aes(log2FoldChange, Family)) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

>>>>

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- loadQiimeData()

@@

gg_topi.family <- tax_glom(gg_topi,taxrank = "Family")
gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
```

@@

### Dotplot log2FC vs Family
```{r}
res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),c(-7,-6)])) %>%
  ggplot(aes(log2FoldChange, Family)) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### 2019-01-06 gg_topi.Rmd
new figures. Fixed family levels boxplots

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])

sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

unass_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(Domain == 'Unassigned') %>%
                select(OTU)

unknown_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Genus)) %>%
                select(OTU)

tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'

@@

gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))

meta_samples <- data.frame(otu_table(gg_topi), check.names = FALSE) %>% 
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
  dplyr::group_by(timepoint, OTU) %>%
  dplyr::select(-SampleID)%>%
  dplyr::summarise(Abundance=mean(Abundance)) %>%
  tidyr::spread(timepoint, Abundance) %>%
  tibble::column_to_rownames(var = 'OTU') %>%
  data.frame()

gg_topi_meta_samples <- phyloseq(otu_table(meta_samples, taxa_are_rows = TRUE), tax_table(gg_topi), phy_tree(gg_topi))
```
<!-- # Taxonomy profile at genus level -->
<!-- ```{r taxonomy, cache=FALSE} -->
<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
<!-- ``` -->

@@

# Alpha diversity
```{r Alpha diversity}
# otu_table(gg_topi) %>% colSums() %>% sort %>% barplot(xlab = c())
# otu_table(gg_topi) %>% colSums() %>% summary()

alpha_div <- estimate_richness(gg_topi)
alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))

x <- melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon'))
x$value <- as.numeric(x$value)

ggplot(x) +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~condizione,scales = 'free_y')
```
<!-- ## SFB abundance -->
<!-- ```{r} -->
<!-- knitr::kable(otu_table(gg_topi)[c(879,880),]) -->
<!-- ``` -->

# Beta diversity
```{r Beta diversity, cache=FALSE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")
distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")

ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordwu_meta <-  ordinate(gg_topi_meta_samples, method = "MDS", distance = distwu_meta, weighted = TRUE)
ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

# plot_ordination(gg_topi, ordwu, color = 'condizione',shape = 'timepoint', title = 'Weighted UniFrac')
# plot_ordination(gg_topi, ordwu, color = 'merge_timepoint', title = 'Weighted UniFrac')

distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesbc <- reshape2::melt(as.matrix(distbc), varnames = c('rows','cols'))
distancesbc$condizione1 <- sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][2])
distancesbc$condizione2 <- sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][2])
distancesbc$topo1 <- factor(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$topo2 <- factor(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$timepoint1 <- stringr::str_replace(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesbc$timepoint2 <- stringr::str_replace(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
```

@@

subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=tumor_injected), size=3, alpha=.8) +
  scale_color_brewer(name = "Tumor injected", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = -0.2, xmax = -.2, ymin = .25, ymax = .25)
```

## Alpha diversity: tumor vs no tumor
```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T2','T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2)))

ttest_res <- data.frame()
for(measure in c('Observed', 'se.chao1', 'Simpson', 'Shannon')){
  ttest_res %<>% bind_rows(data.frame(Measure=measure, pvalue=with(filt_alpha, t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
}

filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~.,scales = 'free_y')

@@

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances_tmp <- distances
distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
distances_tmp[distances_tmp$timepoint2 == 1 | distances_tmp$timepoint2 == 2, 'timepoint2'] <- 2

@@

## Beta diversity between timepoints, T(-1) is the baseline
```{r Beta diversity between timepoints tn-1}
tn1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==timepoint2-1) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

@@

## Beta diversity T3 vs T4
```{r}
t1 <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==4) %>%
  dplyr::arrange(topo1)

@@

## Beta diversity T3 vs T7
```{r}
t37 <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==7) %>%
  dplyr::arrange(topo1)

@@

## Correlation beta-diversity T3-TX, tumor mass
```{r corr tumor wunifrac}
distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

# Distribution of distances between conditions per timepoint
```{r}
dct <- distances %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```

@@

# Distribution of distances between timepoints per conditions
```{r}
dtc <- distances %>%
  filter(condizione1 == condizione2) %>%
  dplyr::rename(condizione = condizione1) %>% 
  group_by(condizione, timepoint1, timepoint2) %>%
  summarise(mediandistance=median(value)) 
```

>>>>

source('R/loadQiimeData.R')
source('R/calculatePairwiseDistances.R')
```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- loadQiimeData()

@@

gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))
```
<!-- # Taxonomy profile at genus level -->
<!-- ```{r taxonomy, cache=FALSE} -->
<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
<!-- ``` -->

@@

```{r Calculate alpha diversity}
alpha_div <- estimate_richness(gg_topi)
alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))
```

# Alpha diversity: timepoint x Index x Condition
```{r Plot Alpha diversity boxplots}
melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~condizione,scales = 'free_y')
```

## Alpha diversity: tumor vs no tumor
```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T1', 'T2', 'T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  rename('se.chao1' = 'Chao1') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2))) %>% droplevels()

ttest_res <- data.frame()
for ( ts in list(c('T1','T2'), c('T2','T3')) ){
  for( measure in c('Observed', 'Chao1', 'Simpson', 'Shannon') ){
    ttest_res %<>% bind_rows( data.frame(Measure = measure,
                                         Timepoint = paste(ts, collapse = ' vs '),
                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
  }
}

filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  xlab('Timepoint') +
  ylab('Value') +
  facet_grid(variable~.,scales = 'free_y') +
  theme_minimal() +
  theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.y = element_text(angle = 0))

# Beta diversity
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")

ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

distanceswu <- calculate_pairwise_distances(distwu)
distancesbc <- calculate_pairwise_distances(distbc)
distancesjac <- calculate_pairwise_distances(distjac)
```

@@

subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
```

@@

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances_tmp <- distanceswu
distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
distances_tmp[distances_tmp$timepoint2 == 1 | distances_tmp$timepoint2 == 2, 'timepoint2'] <- 2

@@

## Beta diversity between timepoints, T(-1) is the baseline
```{r Beta diversity between timepoints tn-1}
tn1 <- distanceswu %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==timepoint2-1) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

@@

## Beta diversity T3 vs T4
```{r}
t1 <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==4) %>%
  dplyr::arrange(topo1)

@@

## Beta diversity T3 vs T7
```{r}
t37 <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==7) %>%
  dplyr::arrange(topo1)

@@

## Correlation beta-diversity T3-TX, tumor mass
```{r corr tumor wunifrac}
distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

# Distribution of distances between conditions per timepoint
```{r}
dct <- distanceswu %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```

@@

# Distribution of distances between timepoints per conditions
```{r}
dtc <- distanceswu %>%
  filter(condizione1 == condizione2) %>%
  dplyr::rename(condizione = condizione1) %>% 
  group_by(condizione, timepoint1, timepoint2) %>%
  summarise(mediandistance=median(value)) 
```

### 2019-01-06 paper_figures.Rmd
new figures. Fixed family levels boxplots

```{r QIIME data loading}
gg_topi <- loadQiimeData()
```

@@

## Panel C
```{r}

>>>>

```{r QIIME data loading}
gg_topi <- loadQiimeData()
gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.family <- tax_glom(gg_topi,taxrank = "Family")
gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
```

@@

## Panel C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.
```{r DESeq2 loading, cache=TRUE}
pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)
res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt_genus <- res_genus[!is.na(res_genus$padj),]
res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]

t23_samples_idx_genus <- grep('T2|T3', sample_names(gg_topi.genus.relab), perl = TRUE)

genus_ab_t <- otu_table(gg_topi.genus.relab)[rownames(res.filt_genus), t23_samples_idx_genus] %>%
            data.frame %>%
            tibble::rownames_to_column("OTU") %>%
            tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
            mutate(timepoint=factor(stringr::str_sub(.$SampleID,12, 13)))
```

```{r}
lacno_bact_relabs <-genus_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>% filter(Family %in% c('Bacteroidales S24-7 group') | Genus %in% c('Lachnospiraceae NK4A136 group'))

lacno_bact_relabs %>% group_by(timepoint, interaction(Family, Genus)) %>% dplyr::summarise(Mean=mean(Abundance), SD = sd(Abundance)) %>% knitr::kable()

lacno_bact_relabs %>% 
  mutate(Taxonomy=interaction(Family, Genus)) %>%
  select(Abundance, Taxonomy, timepoint) %>%
  group_by(timepoint, Taxonomy) %>%
  summarise(Abundance = list(Abundance)) %>% 
  spread(timepoint, Abundance) %>% 
  group_by(Taxonomy) %>%
  dplyr::summarise("Wilcoxon p value"=wilcox.test(unlist(T2), unlist(T3), paired = TRUE)$p.value)  %>% knitr::kable()
```

```{r}
  lacno_bact_relabs %>%
  ggplot() +
    geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
    facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
    theme_minimal() +
    guides(fill=FALSE) +
    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.x = element_text(size=12))
